Reproducible Research: Peer Assessment 1
===============

Download data and process date column

```{r}
require(plyr)
require(Hmisc)
df <- read.csv("./activity.csv")
df$date <- as.Date(as.character(df$date))
df$date <- as.factor(df$date)
str(df)
```

####What is mean total number of steps taken per day?

Calculate total steps per day
```{r echo=TRUE}
day_total <- aggregate(df$steps, list(date = df$date), sum)
print(day_total)
```

Calculate mean and median
```{r}
mean(day_total$x, na.rm=TRUE)
```
```{r}
median(day_total$x, na.rm=TRUE)
```

Make a plot 
(Note:  I've elected to product a bar plot becuase I think it communicates the data clearer)
```{r echo=TRUE}
names(day_total)[2] <- "total_steps"
barplot(day_total$total_steps, names.arg=day_total$date, ylab="Steps", xlab="Date", col="blue")
```

####What is the average daily activity pattern?
Calculate and plot
```{r}
interval_steps <- aggregate(x=list(steps=df$steps), 
                      by=list(interval=df$interval), 
                      FUN=mean, na.rm=TRUE)
plot(interval_steps$interval, interval_steps$steps, xlab="Interval", ylab="Steps", type = "l", col="red")
```

Max 5-min interval
```{r}
interval_steps[which.max(interval_steps[,2]), ]
```

####Imputing missing values
Calculate the total number of NA
```{r}
sum(is.na(df$steps))
```

Strategy for filling-in NA values.  I've decided to replace each NA with the mean of the entire steps column.
```{r}
df_mean <- mean(df$steps, na.rm = TRUE)
df_no_na <- df
df_no_na[is.na(df_no_na)] <- df_mean
head(df_no_na, 20)
```
Verify there are no NAs left
```{r}
sum(is.na(df_no_na$steps))
```
Make a plot (same barplot explanation as above)
```{r}
no_na_total <- aggregate(df_no_na$steps, list(date = df_no_na$date), sum)
names(no_na_total)[2] <- "total_steps"
barplot(no_na_total$total_steps, names.arg=no_na_total$date, ylab="Steps", xlab="Date", col="blue")
```
Calculate mean and median
```{r}
mean(no_na_total$total_steps, na.rm=TRUE)
```
```{r}
median(no_na_total$total_steps, na.rm=TRUE)
```
What effect does imputing have?  
Imputing the mean into every NA leaves the mean unchanged but changes the median slightly.



####Are there differences in activity patterns between weekdays and weekends?
```{r}

df3 <- mutate(df, day = weekdays(as.Date(df$date)))
head(df3)
tail(df3)
```
Subset out Saturday and Sunday
```{r}
df4 <- subset(df3, day == "Saturday")
df5 <- subset(df3, day == "Sunday")
df6 <- rbind(df4, df5)
head(df6)
tail(df6)
```
Plot the weekend data in 5-min time series
```{r}
weekend_data <- aggregate(x=list(steps=df6$steps), 
                      by=list(interval=df6$interval), 
                      FUN=mean, na.rm=TRUE)
plot(weekend_data$interval, weekend_data$steps, xlab="Weekend Interval", ylab="Steps", type = "l", col="red")
```

plot weekday data
```{r}
df44 <- subset(df3, day == "Monday")
df45 <- subset(df3, day == "Tuesday")
df46 <- subset(df3, day == "Wednesday")
df47 <- subset(df3, day == "Thursday")
df48 <- subset(df3, day == "Friday")
df6 <- rbind(df44, df45, df46, df47, df48)
head(df6)
tail(df6)
```
Plot weekday data
```{r}
weekday_data <- aggregate(x=list(steps=df6$steps), 
                      by=list(interval=df6$interval), 
                      FUN=mean, na.rm=TRUE)
plot(weekday_data$interval, weekday_data$steps, xlab="Weekday Interval", ylab="Steps", type = "l", col="blue")
```
